---
title: "Exploring Tweets with Wordclouds in R"
author: "Deepal DSilva"
date: "April 13, 2018"
output: html_document
---

```{r setup, include=FALSE, message=F, warning=F}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

##It's easy! You just need to use the WordCloud2 R package
 
Ever wondered what you tweet about the most? Or you have your favourite conference going on and you want to know what's the buzz around it?
Or maybe you just want to know what's the talk about the latest movie that got released?

Well WordClouds are your friend. They are simple to set up, provide stunning visualizations and easily customizable. 

##But first, what is a WordCloud?

An image composed of words used in a particular text or subject, in which the size of each word indicates its frequency or importance.

Now that you know the basics, let's get started.

##Loading up the required libraries

```{r}
library(twitteR)
library(ROAuth)
library(stringr)
library(tm)
library(wordcloud2)
library(tidytext)
```

##Twitter App set up

We are going to use twitter data to build our wordcloud, so go get a twitter account if you do not have one. I'll wait...

Next we are going to need a Twitter App.

This just a one time setup. Think of it as:

  1. You need to authenticate yourself on Twitter
  
  2. So that you can send a request for tweets and
  
  3. For twitter to send them back to you.
  
I'll not go into the detailed steps. You can use one of these tutorials available. [Read](https://iag.me/socialmedia/how-to-create-a-twitter-app-in-8-easy-steps/) or [Watch](https://www.youtube.com/watch?v=xqSp7060Gj0)


```{r}
consumer_key <- "mftPn6qtOfCL46fzJY3pTeVI4"   #Your Consumer Key (API Key)
consumer_secret <- "x5D1VIlV1O4TkZPshj31j88CutI8X2eSfgryy0pdMpDJjUTcKK"   #Your Consumer Secret (API Secret)
access_token <- "2924790128-WORx3u9bTwDqAYr6zSanDnDlafLaSHBdzVKYZxC"   # Your Access Token
access_secret <- "aMFTRClgP3Q2K0pJ21p1fromq1OWx4cS1hU6i4qJyqpZb"    #Your Access Token Secret
setup_twitter_oauth(consumer_key, consumer_secret, access_token, access_secret)

#These tokens will not work. You need to enter in your own. 
```

  
##Extracting Twitter Data
 
You can either extract tweets based on a user's profile or any keywords/hashtags. Here are both the examples.

```{r}
#Query a hashtag
#tweets <- searchTwitter("#rstats",n=3000,lang="en", resultType = "popular")

#OR 

#Query a user you follow or yourself
tweets <- userTimeline("dsilvadeepal",n=3200,includeRts = FALSE)
```

##Text Mining your tweets

We now need to extract the text from the tweets to a vector.

```{r}
tweets.txt <- sapply(tweets, function(t)t$getText())
```

We are first going to remove graphical parameters. This removes visible characters (anything except spaces and control characters) to avoid input errors.

```{r}
tweets.txt <- str_replace_all(tweets.txt,"[^[:graph:]]", " ") 
```

Now let's create a function to clean out tweets. Here we are going to remove digits, punctuations, spaces, http links, retweets (RTs). You can customize this function based on the data you are processing


```{r}
clean.text = function(x)
{
  
  x = tolower(x)                   # tolower
  x = gsub("rt", "", x)            # remove rt
  x = gsub("@\\w+", "", x)         # remove at
  x = gsub("[[:punct:]]", "", x)   # remove punctuation
  x = gsub("[[:digit:]]", "", x)   # remove numbers
  x = gsub("http\\w+", "", x)      # remove links http
  x = gsub("[ |\t]{2,}", "", x)    # remove tabs
  x = gsub("^ ", "", x)            # remove blank spaces at the beginning
  x = gsub(" $", "", x)            # remove blank spaces at the end
  
  return(x)
}

```

Now let's call our function on the tweets we have. 

```{r}
clean_tweet <- clean.text(tweets.txt)
```

Next we build a Corpus - A Corpus is a collection of text documents and the VectorSource points to the vector where the tweets are stored.


```{r}
tweets <- Corpus(VectorSource(clean_tweet))
```

We are also going to create a vector to remove english stop words and any additional words that are irrelevant (we'll identify these words in a later step). Some common english stop words are 'the', 'I', 'he'. 
```{r}
wordsToRemove <- c(stopwords('en'), 'tco', 'https')
clean_tweet <- tm_map(tweets, removeWords, wordsToRemove)
```


We now create a Term Document Matrix (TDM) which reflects the number of times each word in the corpus is found in each of the documents.

```{r}
dtm <- TermDocumentMatrix(clean_tweet, control = list(wordLengths = c(1, Inf)))
m <- as.matrix(dtm)
v <- sort(rowSums(m),decreasing=TRUE)
d <- data.frame(word = names(v),freq=v)
head(d, 10)   #inspect our word list and remove any irrelevant words in the stopwords step above
```

Finally for the fun part!

##Creating our wordcloud

Here we are going to use the WordCloud2 package which is a newer package with a lot more customizations.

```{r}
wordcloud2(d, shape = "star", color="random-dark", backgroundColor = "lightblue",
           minRotation = -pi/4, maxRotation = -pi/4, size = 0.5)

```

There are a few limitations to note using the TwitterR package:
  1. Twitter API limits the number of tweets a user can get from a particular timeline to be 3200

