Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:

***

### Scatterplots
Notes:

```{r Scatterplots}
library(ggplot2)
pf <- read.csv("pseudo_facebook.tsv", sep = "\t")

qplot(age, friend_count, data = pf)
```

***

#### What are some things that you notice right away?
Response:

Younger people have much higher friend counts. Few vertical bars around 69, 100 age with high friend counts..which is most likely fake ages..
***

### ggplot Syntax
Notes:

```{r ggplot Syntax}
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_point(alpha = 1/20) +
  xlim(13, 90)

summary(pf$age)
```

***

### Overplotting
Notes:

```{r Overplotting}
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_jitter(alpha = 1/20) +
  xlim(13, 90)
```

#### What do you notice in the plot?
Response:

***

### Coord_trans()
Notes:

```{r Coord_trans()}
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_jitter(alpha = 1/20, position = position_jitter(height = 0)) +
  xlim(13, 90) +
  coord_trans(y = "sqrt") 
```

***

### Alpha and Jitter
Notes:

```{r Alpha and Jitter}
ggplot(aes(x = age, y = friendships_initiated), data = pf) +
  geom_jitter(alpha = 1/30, position = position_jitter(height = 0)) +
  coord_trans(y = "sqrt") +
  xlim(13,90)
```

***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:

```{r Conditional Means}
library(dplyr)

age_groups <- group_by(pf, age)

pf.fc_by_age <- summarise(age_groups,
          friend_count_mean = mean(friend_count),
          friend_count_median = median(friend_count),
          n = n())

pf.fc_by_age <- arrange(pf.fc_by_age)

head(pf.fc_by_age)
```
#Conditional means alternate plot

```{r}
library("dplyr")

pf.fc_by_age <- pf %>%
  group_by(age) %>%
  summarise(friend_count_mean = mean(friend_count),
          friend_count_median = median(friend_count),
          n = n()) %>%
  arrange(age)

head(pf.fc_by_age, 20)
```
 
Create your plot!

```{r Conditional Means Plot}
ggplot(aes(x = age, y = friend_count_mean), data = pf.fc_by_age) +
  geom_line()

```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_jitter(alpha = 1/20,
              position = position_jitter(height = 0),
              color = "orange") +
  xlim(13, 90) +
  coord_trans(y = "sqrt") +
  geom_line(stat = "summary", fun.y = mean) +
  geom_line(stat = "summary", fun.y = quantile, fun.args = list(probs = 0.1), linetype = 2, color = "blue") +
  geom_line(stat = "summary", fun.y = quantile, fun.args = list(probs = 0.9), linetype = 2, color = "blue") +
  geom_line(stat = "summary", fun.y = quantile, fun.args = list(probs = 0.5), color = "blue")
```


```{r}
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_jitter(alpha = 1/20,
              position = position_jitter(height = 0),
              color = "orange") +
  coord_cartesian(xlim = c(13, 70), ylim = c(0, 1000)) +
  geom_line(stat = "summary", fun.y = mean) +
  geom_line(stat = "summary", fun.y = quantile, fun.args = list(probs = 0.1), linetype = 2, color = "blue") +
  geom_line(stat = "summary", fun.y = quantile, fun.args = list(probs = 0.9), linetype = 2, color = "blue") +
  geom_line(stat = "summary", fun.y = quantile, fun.args = list(probs = 0.5), color = "blue")
```

#### What are some of your observations of the plot?
Response:

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}
cor.test(pf$age, pf$friend_count, method = "pearson")

#OR 

with(pf, cor.test(age, friend_count, method = "pearson"))
```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(subset(pf, age <= 70), cor.test(age, friend_count))
```

***

### Correlation Methods
Notes:

***

## Create Scatterplots
Notes:

```{r}
ggplot(aes(x = www_likes_received, y = likes_received), data = pf) +
  geom_point()
```

***

### Strong Correlations
Notes:

```{r Strong Correlations}
ggplot(aes(x = www_likes_received, y = likes_received), data = pf) +
  geom_point() +
  xlim(0, quantile(pf$www_likes_received, 0.95)) +
  ylim(0, quantile(pf$likes_received, 0.95)) +
  geom_smooth(method = "lm", color = "red")
```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}
cor.test(pf$www_likes_received, pf$likes_received)
```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
#install.packages('alr3')
library(alr3)
data(Mitchell)
?Mitchell

```

Create your plot!

```{r Temp vs Month}
ggplot(aes(x = Month, y = Temp), data = Mitchell) + 
  geom_point()
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}
cor.test(Mitchell$Month, Mitchell$Temp)
```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}
range(Mitchell$Month)

ggplot(aes(x = Month, y = Temp), data = Mitchell) + 
  geom_point() +
  scale_x_continuous(breaks = seq(0, 203, 12))
```

***

### A New Perspective

What do you notice?
Response:

```{r}
ggplot(aes(x=(Month%%12),y=Temp), data=Mitchell)+
  geom_point()
```

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}
ggplot(aes(x = age, y = friend_count_mean), data = pf.fc_by_age) +
  geom_line()

head(pf.fc_by_age, 10)
pf.fc_by_age[17:19,]
```

***

### Age with Months Means

```{r Age with Months Means}
pf$age_with_months <- pf$age + (1 - pf$dob_month/12)
head(pf$age_with_months)
```

Programming Assignment
```{r Programming Assignment}

library(dplyr)

pf.fc_by_age_months <- pf %>%
  group_by(age_with_months) %>%
  summarise(friend_count_mean = mean(friend_count),
          friend_count_median = median(friend_count),
          n = n()) %>%
  arrange(age_with_months)

head(pf.fc_by_age_months)
```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}
ggplot(aes(x = age_with_months, y = friend_count_mean), 
       data = subset(pf.fc_by_age_months, age_with_months < 71)) +
  geom_line()
```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}
p1 <- ggplot(aes(x = age, y = friend_count_mean), 
       data = subset(pf.fc_by_age, age < 71)) +
  geom_line() +
  geom_smooth()


p2 <- ggplot(aes(x = age_with_months, y = friend_count_mean), 
       data = subset(pf.fc_by_age_months, age_with_months < 71)) +
  geom_line() +
  geom_smooth()

p3 <- ggplot(aes(x = round(age/5)*5, y = friend_count), data = subset(pf, age < 71)) +
  geom_line(stat = "summary", fun.y = mean)

library(gridExtra)
grid.arrange(p2, p1, p3, ncol = 1)
```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

