---
title: "Lesson 6 Problem Set - Explore 2 Variables"
author: "Deepal DSilva"
date: "April 9, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load the diamonds dataset
```{r}
library(ggplot2)
library(dplyr)
data("diamonds")
summary(diamonds)
```

Create a scatterplot of price vs x. using the ggplot syntax.

```{r}
ggplot(aes(x = price, y = x), data = diamonds) +
  geom_point()
```
Correlations

```{r}
cor.test(diamonds$price, diamonds$x)

cor.test(diamonds$price, diamonds$y)

cor.test(diamonds$price, diamonds$z)
```

Create a simple scatter plot of price vs depth.

```{r}
ggplot(aes(x = price, y = depth), data = diamonds) +
  geom_point(alpha = 1/100) +
  scale_x_continuous(breaks = seq(0,10,2))
```

Correlations of depth vs price

```{r}
cor.test(diamonds$depth, diamonds$price)
```

Create a scatterplot of price vs carat and omit the top 1% of price and carat values.

```{r}
ggplot(aes(x = carat, y = price), data = diamonds) +
  geom_point() +
  xlim(0, quantile(diamonds$carat, 0.99)) +
  ylim(0, quantile(diamonds$price, 0.99))
```


Create a scatterplot of price vs. volume (x * y * z). This is a very rough approximation for a diamond's volume.

Create a new variable for volume in the diamonds data frame.

```{r}
diamonds$volume <- diamonds$x * diamonds$y * diamonds$z

head(diamonds)

ggplot(aes(x = volume, y = price), data = diamonds) + 
  geom_point()
```

Count diamonds with volume = 0

```{r}
summary(diamonds$volume == 0)
```

What is the correlation between price and volume
Exclude diamonds with 0 volums or greater than or equal to 800

```{r}
new_diamonds <- subset(diamonds, diamonds$volume > 0 & diamonds$volume < 800)
head(new_diamonds)

cor.test(new_diamonds$price, new_diamonds$volume)

```

Subset the data to exclude diamonds with a volume greater than or equal to 800. Also, exclude diamonds with a volume of 0. Adjust the transparency of the points and add a linear model to the plot.

```{r}
ggplot(aes(x = volume, y = price), data = new_diamonds) +
  geom_point(alpha = 1/50, color = "orange") + 
  geom_smooth(method = lm)
```
Use the function dplyr package to create a new data frame containing info on diamonds by clarity.

Name the data frame diamondsByClarity. The data frame should contain the following variables in this order.

       (1) mean_price
       (2) median_price
       (3) min_price
       (4) max_price
       (5) n

 where n is the number of diamonds in each level of clarity.

```{r}
diamondsByClarity <- diamonds %>%
  group_by(clarity) %>%
  summarise(mean_price = mean(price),
          median_price = median(price),
          min_price = min(price),
          max_price = max(price),
          n = n())

head(diamondsByClarity)

```


```{r}
library(gridExtra)

diamonds_by_clarity <- group_by(diamonds, clarity)
diamonds_mp_by_clarity <- summarise(diamonds_by_clarity, mean_price = mean(price))

diamonds_by_color <- group_by(diamonds, color)
diamonds_mp_by_color <- summarise(diamonds_by_color, mean_price = mean(price))

a <- ggplot(aes(x = clarity, y = mean_price), data = diamonds_mp_by_clarity) + 
  geom_bar(stat = "identity")

b <- ggplot(aes(x = color, y = mean_price), data = diamonds_mp_by_color) + 
  geom_bar(stat = "identity")

grid.arrange(a, b, ncol = 1)
```

