Lesson 3
========================================================

***

### What to Do First?
Notes:
Check working dir and change to the correct dir.
```{r}
getwd()
list.files()
```

***

### Pseudo-Facebook User Data
Notes: Read the data. Remember it is a tab seperated values file.

You can also use read.delim('pseudo_facebook.tsv') . The read.delim() function defaults to the tab character as the separator between values and the period as the decimal character

```{r Pseudo-Facebook User Data}
pf <-  read.csv("pseudo_facebook.tsv", sep = "\t")

```

***

### Histogram of Users' Birthdays
Notes:
You can also use 
ggplot(aes(x = dob_day), data = pf) +
  geom_histogram(binwidth = 1) +
  scale_x_continuous(breaks = 1:31)

```{r Histogram of Users\' Birthdays}
#install.packages('ggplot2')
library(ggplot2)
theme_set(theme_minimal(24)) 

names(pf)
qplot(x = dob_day, data = pf) +
  scale_x_continuous(breaks = 1:31) +
  facet_wrap(~dob_month, ncol = 3)

```

***

#### What are some things that you notice about this histogram?
Response:

***

### Moira's Investigation
Notes:

***

### Estimating Your Audience Size
Notes:

***

#### Think about a time when you posted a specific message or shared a photo on Facebook. What was it?
Response:

#### How many of your friends do you think saw that post?
Response:

#### Think about what percent of your friends on Facebook see any posts or comments that you make in a month. What percent do you think that is?
Response:

***

### Perceived Audience Size
Notes:

***
### Faceting
Notes:

```{r Faceting}

```

#### Let’s take another look at our plot. What stands out to you here?
Response:

***

### Be Skeptical - Outliers and Anomalies
Notes:

***

### Moira's Outlier
Notes:
#### Which case do you think applies to Moira’s outlier?
Response:

***

### Friend Count
Notes:

#### What code would you enter to create a histogram of friend counts?

```{r Friend Count}
qplot(x=friend_count, data = pf)
```

#### How is this plot similar to Moira's first plot?
Response:

***

### Limiting the Axes
Notes:

```{r Limiting the Axes}
qplot(x=friend_count, data = pf) +
  scale_x_continuous(limits = c(0,1000))

```

### Exploring with Bin Width
Notes:

***

### Adjusting the Bin Width
Notes:

### Faceting Friend Count
```{r Faceting Friend Count}
# What code would you add to create a facet the histogram by gender?
# Add it to the code below.
qplot(x = friend_count, data = pf, binwidth = 10) +
  scale_x_continuous(lim = c(0, 1000), breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
```

***

### Omitting NA Values
Notes:

```{r Omitting NA Values}
qplot(x = friend_count, data = subset(pf, !is.na(gender)), binwidth = 10) +
  scale_x_continuous(lim = c(0, 1000), breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
```

***

### Statistics 'by' Gender
Notes:

```{r Statistics \'by\' Gender}
table(pf$gender)
```

#### Who on average has more friends: men or women?
Response:
```{r}
by(pf$friend_count, pf$gender, summary)
```

#### What's the difference between the median friend count for women and men?
Response:

#### Why would the median be a better measure than the mean?
Response:

***

### Tenure
Notes:
Using ggplot 

ggplot(aes(x = tenure), data = pf) +
  geom_histogram(binwidth = 30, color = 'black', fill = '#099DD9')

ggplot(aes(x = tenure/365), data = pf) +
  geom_histogram(binwidth = .25, color = 'black', fill = '#F79420')
  
```{r Tenure}
qplot(x = tenure, data = pf, binwidth = 30,
      color = I("black"), fill = I("#099099"))
  
```

***

#### How would you create a histogram of tenure by year?

```{r Tenure Histogram by Year}
qplot(tenure/365, data=pf, binwidth= .25,
      color = I("black"), fill = I("#69AFD9"))+
  scale_x_continuous(breaks = seq(1,7,1), limits = c(0,7))
```

***

### Labeling Plots
Notes:

```{r Labeling Plots}
qplot(tenure/365, data=pf, binwidth= .25,
      xlab = "Number of years using Facebook",
      ylab = "Number of users in sample",
      color = I("black"), fill = I("#69AFD9"))+
  scale_x_continuous(breaks = seq(1,7,1), limits = c(0,7))
```

***

### User Ages
Notes:

```{r User Ages}
qplot(x = age, data=pf, binwidth = 1,
      xlab = "Age of Facebook users",
      ylab = "Number of users in sample",
      color = I("black"), fill = I("#69AFD9")) +
  scale_x_continuous(breaks = seq(0,120,5))
```

#### What do you notice?
Response:

***

### The Spread of Memes
Notes:

***

### Lada's Money Bag Meme
Notes:

***

### Transforming Data
Notes:
```{r}
qplot(x=friend_count, data=pf)

summary(pf$friend_count)

summary(log10(pf$friend_count +1))

summary(sqrt(pf$friend_count))
```

***

### Add a Scaling Layer
Notes:

ggplot(aes(x = friend_count), data = pf) +
  scale_x_log10()

```{r Add a Scaling Layer}
library(gridExtra)

p1 <-  qplot(x=friend_count, data=pf)
p2 <-  qplot(x=log10(pf$friend_count + 1), data=pf)
p3 <-  qplot(x=sqrt(pf$friend_count), data=pf)
grid.arrange(p1,p2,p3, ncol=1)
```

***


### Frequency Polygons

```{r Frequency Polygons}
qplot(x = friend_count,y = ..count../sum(..count..),
      data = subset(pf, !is.na(gender)),
      binwidth = 10, geom = "freqpoly", color = gender) +
  scale_x_continuous(lim = c(0, 1000), breaks = seq(0, 1000, 50)) 
```

***

### Likes on the Web
Notes:

```{r Likes on the Web}
qplot(x = www_likes,
      data = subset(pf, !is.na(gender)),
      geom = "freqpoly", color = gender) +
  scale_x_continuous() +
  scale_x_log10()
```

```{r}
by(pf$www_likes, pf$gender, sum)
```

***

### Box Plots
Notes:

```{r Box Plots}
qplot(x = gender, y = friend_count, 
      data = subset(pf, !is.na(gender)), 
      geom = "boxplot")


```

#### Adjust the code to focus on users who have friend counts between 0 and 1000.

```{r}
qplot(x = gender, y = friend_count, 
      data = subset(pf, !is.na(gender)), 
      geom = "boxplot", ylim = c(0,1000))

qplot(x = gender, y = friend_count, 
      data = subset(pf, !is.na(gender)), 
      geom = "boxplot") +
  coord_cartesian(ylim = c(0,1000))
```

***

### Box Plots, Quartiles, and Friendships
Notes:

```{r Box Plots, Quartiles, and Friendships}
qplot(x = gender, y = friend_count, 
      data = subset(pf, !is.na(gender)), 
      geom = "boxplot") +
  coord_cartesian(ylim = c(0,250))

by(pf$friend_count, pf$gender, summary)
```

#### On average, who initiated more friendships in our sample: men or women?
Response:
#### Write about some ways that you can verify your answer.
Response:
```{r Friend Requests by Gender}
qplot(x = gender, y = friendships_initiated, 
      data = subset(pf, !is.na(gender)), 
      geom = "boxplot") +
  coord_cartesian(ylim = c(0,150))

by(pf$friendships_initiated, pf$gender, summary)
```

Response:

***

### Getting Logical
Notes:

```{r Getting Logical}
summary(pf$mobile_likes)

summary(pf$mobile_likes > 0)

mobile_check_in <-  NA
pf$mobile_check_in <- ifelse(pf$mobile_likes > 0, 1, 0)
pf$mobile_check_in <- factor(pf$mobile_check_in)
summary(pf$mobile_check_in)
sum(pf$mobile_check_in ==1) / length(pf$mobile_check_in)

```

Response:

***

### Analyzing One Variable
Reflection:

Take a close look at individual variables
Understand they types of values they take on, what the distributions ook like
If they are missing values or are outliers.
Histograms, box plots, frequency polygons
Change limits on axes
Adjust bin width on histograms
Transformed variables logarithmically or  turned them into binary

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!