---
title: "Lesson 4 Problem Set - Explore 1 Variable"
author: "Deepal DSilva"
date: "April 4, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load the diamonds dataset
```{r}
library(ggplot2)
data("diamonds")
summary(diamonds)
```

Look at the colors in the diamond data set
```{r}
diamonds$color
``` 

Create a histogram of the price of all the diamonds in the diamond data set.

```{r}
qplot(x = price, data = diamonds, binwidth = 500, title = "Diamond Price Distribution",
                                                         xlab ="Diamond Price USD", ylab = "Frequency")  

mean(diamonds$price)

median(diamonds$price)
```


Supposed we want to know the following:

How many cost less than U$500?
How many cost less than U$250?
How many cost equal to U$15,000 or more?

```{r}
sum(diamonds$price < 500)
sum(diamonds$price < 250)
sum(diamonds$price >= 15000)
```

Analyse the cheaper diamonds

```{r}
qplot(x = price, data = diamonds, binwidth = 10, title = "Diamond Price Distribution",
      xlab ="Diamond Price USD", ylab = "Frequency")  +
      scale_x_continuous(limits = c(0,2500))   

ggsave('priceHistogram.png')
```
  Break out the histogram of diamond prices by cut
  
```{r}
qplot(x = price, data = diamonds, binwidth = 50, title = "Diamond Price Distribution By Cut",
      xlab ="Diamond Price USD", ylab = "Frequency")  +
      scale_x_continuous(limits = c(0,2500))   +
  facet_wrap(~cut)
```
  
                                                         
Which cut has the highest priced diamond?
Premium

Which cut has the lowest priced diamond?
Premium & Ideal

Which cut has the lowest median price?
Ideal

```{r}
by(diamonds$price, diamonds$cut, max)

by(diamonds$price, diamonds$cut, min)

by(diamonds$price, diamonds$cut, median)

```

Change the y-axis to be different for each histogram. If the scales are fixed you might miss some patterns in the data
```{r}
qplot(x = price, data = diamonds, binwidth = 100, title = "Diamond Price Distribution By Cut",
      xlab ="Diamond Price USD", ylab = "Frequency")  +
      scale_x_continuous()   +
  facet_wrap(~cut, scales = "free_y")
```

Create a histogram of price per carat and facet it by cut. Adjust the bin width and transform the scale of the x-axis using log10.

```{r}
qplot(x = price/carat, data= diamonds, binwidth = .01) +
  scale_x_log10() +
  facet_wrap(~cut, scales="free")
```

Investigate the price of diamonds using box plots, numerical summaries, and one of the following categorical variables: cut, clarity, or color.

```{r}
qplot(x = clarity, y = price, data = diamonds, title = "Diamond Price by Clarity", 
      xlab ="Clarity of Diamond", ylab = "Diamond Price USD", geom = "boxplot") + 
  scale_y_continuous(limits = c(0, 10000))

ggsave('priceBoxplot.png')
```

